---
title: "testpoolmpile"
author: "Casey Flay"
date: "30/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RLinuxModules)
library(data.table)
library(glue)
library(here)
library(tidyverse)
library(stringr)
library(ggplot2)
library(magrittr)
module("load openlava asub bwa/0.7.12 samtools/1.9 bcftools/1.10.2 vcftools")
```

```{r}
mainDir <- "/powerplant/workspace/hrtcdf/FSTs/"
subDir <- "malepool"
royDir <- "/workspace/hrards/projects/hrtcdf/10.alignments/bwa"
outDir <- str_c(mainDir, subDir, sep = "", collapse = TRUE)
outDir

if (dir.exists(mainDir)){
    setwd(file.path(mainDir, subDir))
} else {
    dir.create(file.path(mainDir, subDir))
    setwd(file.path(mainDir, subDir))
}

malepoolfiles <- dir(royDir, pattern = "^HW5YYDSXX", full.names = TRUE, include.dirs = TRUE)

file.symlink (tpoolfiles, outDir)
## we don't need the full pathname into the symlink file.
```

```{r transform filenames to sample, lane and suffix for data.table from link}
wd <- ("/powerplant/workspace/hrtcdf/FSTs/malepool/")
pattern <- c(".bai", ".txt", ".Rmd")
f <- grep(paste(pattern, collapse = "|"), list.files(path="."),  invert=TRUE, value=TRUE)
str1 <- str_replace(f, "XX_" , "XX_&")
str2 <- str_replace(str1, "_L" , "&_L")
str3 <- str_replace(str2, ".bam" , "&.bam&")

malepooldt <- {data.table(str3) %>%
  separate(str3, into = c("HL55", "sample", "lane", "bam", "version"), sep = "&", fill = "left") %>%
  mutate("version" = str_extract(sample, regex("/")))
}
malepooldt

colnames(malepooldt) <- c("HL55", "sample", "lane", "bam", "version" )
#write.table(malepooldt, 'linkpool.tsv')
```

```{bash index.bai}
t=8
#inDir="/powerplant/workspace/hrtcdf/FSTs/TestPool/bamlinkindexed"
logDir="/powerplant/workspace/hrtcdf/FSTs/TestPool/bamlinkindexed/logDir"
if  [ ! -d  "logDir" ]; then mkdir "logDir" ; fi

for vcf in $(ls *.bam); do
bsub -n $t -o $logDir/$vcf.err -e $logDir/$vcf.err "samtools index $vcf > $vcf.bai"
#echo $vcf
done
```

## use samtools for indexing normal?, tabix and stats.
```{bash}
t=8
logDir="/powerplant/workspace/hrtcdf/FSTs/malepool/logDir"
if  [ ! -d  "logDir" ]; then mkdir "logDir" ; fi

for vcf in $(ls *chr25.vcf.gz); do
#bsub -n $t -o $logDir/$vcf.log -e $logDir/$vcf.err "bcftools stats $vcf > ${vcf}.stats "
bsub -n $t -o $logDir/$vcf.log -e $logDir/$vcf.err "bcftools index -t $vcf"
#bsub -n $t -o $logDir/$vcf.log -e $logDir/$vcf.err "samtools index $vcf"
#echo c${vcf}
done
```

```{r run-mileup for chr25}
submit_mpileup <- function(reference = "*.fa.gz", output = c(), bamfiles = c(), job.init = "asub", job.opts = "-q normal") {
  cmd <- "bcftools mpileup -B -Q 0 -f {reference} -r chr25 -a {mpileuptags} {bamfiles} -Ou | bcftools call -f {bcfcalltags} -p {pvalThres} -m -A -Oz -o {output}"
  tmpf <- tempfile(pattern = "submit_mileup.", fileext = ".sh")
  writeLines(text = c(glue(cmd)), con = tmpf)
  submit <- "{job.init} {job.opts} {tmpf}"
  system(glue(submit), intern = TRUE)
}
mpileuptags <- "DP,AD,ADF,ADR,SP,INFO/AD,INFO/ADF,INFO/ADR"
bcfcalltags <- "GQ,GP"
  pvalThres <- 0.99

bamfiles    <- paste(malepooldt$HL55, malepooldt$sample, malepooldt$lane, malepooldt$bam, sep = "")
bamfiles
out         <- paste(malepooldt$sample, "chr25.vcf.gz", sep = "_")
reference   <- dir("/workspace/hrards/projects/hrtcdf/genome/fasta", pattern = ".*\\.gz$", full.names = TRUE)
#file.exists(bamfiles)
#file.exists(reference)
submit_mpileup(reference = reference, output = out, bamfiles = bamfiles)
```

#https://github.com/caseyflay/HighHealth/blob/master/QTLSeqr/CheckAlleleFreqs.Rmd
##Calculate AF over all pools on CHR1
```{r}
bcfcmd <- "bcftools query  -f '%CHROM\t%POS[\t%DP\t%AD]\n' D6_chr25.vcf.gz |awk '{print $1,$2,($4 + $6+ $8 + $10)/($3 + $5  + $7 + $9)}'"
DT <- fread(cmd=bcfcmd,col.names = c('CHR','POS','AF'))
```

```{r message=FALSE, warning=FALSE}
DT[,qplot(AF)]
```

```{bash}
#grep "AF" CK02_01_chr25.bcf.gz
#grep --help
#awk CK02_01_chr25.bcf.gz $1 | $2 'chr25:1737600'
##get header from .gz references: zcat Russell_V2a.fa.gz | grep '>' | less
#view .gz vcf files zcat CK02_01_chr25.bcf.gz | head -n 30

#zcat CK02_01_chr25.bcf.gz | grep -v "386373" | cut -f1,2,3,4,5,6,7,8,9,10 | head -n 30

#zcat CK02_01_chr25.bcf.gz | awk -v "AD" | cut -f1,2,3,4,5,6,7,8,9,10 | head -n 30
```
Andrew Catanach: 
tail -n+2 /powerplant/workspace/cflasc/Kiwifruit/chinensis/20200827_Fisher_tests_for_Y_associated_SNPs/04.significant_p_values/Russell_V2a.chromosomes.and.unassigned_chr25_genotypes_fisher.out \
| sed 's/,/\t/g' | awk '$2>8 {print}' | awk -v OFS="\t" '{print "chr25", $1-1, $1}' 

#See awk '$2>8...., the smaller the number the larger the number of sites returned (and the looser the association) 
#You could copy the file /powerplant/workspace/cflasc/Kiwifruit/chinensis/20200827_Fisher_tests_for_Y_associated_SNPs/04.significant_p_values/Russell_V2a.chromosomes.and.unassigned_chr25_genotypes_fisher.out 
#That will give you all associated sites, then filter for those with the highest -log10_P values
```{bash}
tail -n+2 /powerplant/workspace/cflasc/Kiwifruit/chinensis/20200827_Fisher_tests_for_Y_associated_SNPs/04.significant_p_values/Russell_V2a.chromosomes.and.unassigned_chr25_genotypes_fisher.out \
| sed 's/,/\t/g' | awk '$2>9.3{print}' | awk -v OFS="\t" '{print "chr25", $1-1, $1}' 
```

```{bash}
t=8
logDir="/powerplant/workspace/hrtcdf/FSTs/malepool/logDir"


bsub -n $t -o $logDir/$vcf.log -e $logDir/$vcf.err "zgrep '#CHROM' D6_chr25.vcf.gz > '1chr25_tail.txt' && 
zcat D6_chr25.vcf.gz | tail -n 20 >> '1chr25_tail.txt' "

```

```{bash}
#echo 386374 386412 386444 10841 10851 5186 >> SNP_list.txt 
#zcat input_file.vcf.gz | vcftools --vcf - --site-pi --positions SNP_list.txt --out nucleotide_diversity
#zcat A1_chr25.bcf.gz | vcftools --vcf - --site-pi --positions SNP_list.txt --out malepool
#vcftools --gzvcf A1_chr25.vcf.gz --freq --positions SNP_list.txt --out $vcf
#done
```


#Try make it a bit more bioinformatic
##zcat CKRussell_piled.vcf.gz | head -n 100
##CHROM  POS     ID      REF     ALT     QUAL    FILTER  INFO    FORMAT  D6
##chr1    1       .       C       .       63      .       DP=1;ADF=1;ADR=0;AD=1;MQ0F=0;AN=2;DP4=1,0,0,0;MQ=33     GT:DP:SP:ADF:ADR:AD     0/0:1:0:1:0:1
##[] Format fields must be enclosed in brackets to loop over all samples
#bcfcmd <- "bcftools query  -f '%CHROM\t%POS[\t%DP\t%AD]\n' ../PoolSeq/mpileup/CHR1.cns.vcf.gz| awk '{print $1,$2,($4 + $6+ 

working on this second one, need to pipe to .txt and get correct male%
```{r}
#piledvcf <- paste(lindts$sample, "piled.vcf.gz", sep = "_") -i 'INFO/MAF>0.0001'
bcfcmd <- "bcftools query A1_chr25.vcf.gz -r 'chr25:1-1000000'  -f '[%SAMPLE]\t%CHROM\t%POS\t%REF\t%ALT\t%DP\t%AD\n' | awk '{print $1, $2, $3, $4, $5, $6, $7, ($7+$9+$11+$13), ($8+$10+$12+$14), ($6+$8+$10+$12)/($7+$9+$11+$13), ($7+$9+$11+$13)/$6}'"
DTc <- fread(cmd=bcfcmd, col.names = c('SAMPLE','CHROM','POS','REF','ALT','DP','AD','REFal','ALTal','AF','MaleFreq'))
#DTc %>% write_tsv("test.txt", col_names = TRUE)
#DTc %>% write.table("malefreqDT.txt", append = FALSE, sep = " ", dec = ".", row.names = TRUE, col.names = TRUE)
qplot(POS, MaleFreq, data = DTc, color = DP)
```


```{r}
#DTc[,qplot(POS, MaleFreq)]
#MDT <- 'malefreqDT.txt'
qplot(POS, MaleFreq, data = DTc, color = DP)
#ggplot(data = MDT['POS', 'MaleFreq'], mapping = aes(x= 'POS', y= 'MaleFreq'))

```
###todo need to write to data.frame to capture the data.
Error in `[.data.table`(DTc, POS, MaleFreq) : POS is not found in calling scope but it is a column of type integer. If you wish to select rows where that column contains TRUE, or perhaps that column contains row numbers of itself to select, try DT[(col)], DT[DT$col], or DT[col==TRUE] is particularly clear and is optimized. When the first argument inside DT[...] is a single symbol (e.g. DT[var]), data.table looks for var in calling scope.

##zcat CKRussell_piled.vcf.gz | head -n 100


#git add 'filename'
#git commit 
#git push origin master

