---
title: "Russell alts"
output: html_document
---

**this code will use RussellV2a reference fasta's 3-5kb against Russell2a for haplotigs so that the alternate RussellV2a alleles (haplotigs) can be pulled from the RussellV2a genome .**

use roys zipped reference as reference
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(RLinuxModules)
library(data.table)
library(glue)
library(here)
library(tidyverse)
library(stringr)
library(ggplot2)
library(magrittr)
library(GenomicRanges)
devtools::load_all(here::here())
raw_fastq_directory <- russell_raw_fastq_directory()
raw_fastq_directory
module("load openlava asub/2.2 bwa/0.7.12 samtools/1.9 bcftools/1.10.2 vcftools")
```

```{r, echo = FALSE}
raw_fastq_directory <- "/input/genomic/plant/Actinidia/chinensis/Russell/Genome/Illumina/AGRF_CAGRF12782_C9EUNANXX"

read_pair_from_fastq <- function(file, idx = 7) {
  moniker <- basename(file)
  moniker <- tools::file_path_sans_ext(moniker)
  moniker <- tools::file_path_sans_ext(moniker)
  parts   <- strsplit(moniker, "_")[[1]]
  as.integer(gsub(pattern = "R", replacement = "", x = parts[idx]))
}
lane_from_fastq <- function(file, idx = 6) {
  moniker <- basename(file)
  moniker <- tools::file_path_sans_ext(moniker)
  moniker <- tools::file_path_sans_ext(moniker)
  parts   <- strsplit(moniker, "_")[[1]]
  as.integer(gsub(pattern = "L", replacement = "", x = parts[idx]))
}
barcode_from_fastq <- function(file, idx = 5, delim = "_") {
  moniker <- basename(file)
  moniker <- tools::file_path_sans_ext(moniker)
  moniker <- tools::file_path_sans_ext(moniker)
  parts   <- strsplit(moniker, delim)[[1]]
  parts[idx]
}
sample_name_from_fastq <- function(file, idx = 1, delim = "_") {
  moniker <- basename(file)
  moniker <- tools::file_path_sans_ext(moniker)
  moniker <- tools::file_path_sans_ext(moniker)
  parts   <- strsplit(moniker, delim)[[1]]
  parts[idx]
}
library_size_from_fastq <- function(file, idx = 2, delim = "_") {
  moniker <- basename(file)
  moniker <- tools::file_path_sans_ext(moniker)
  moniker <- tools::file_path_sans_ext(moniker)
  parts   <- strsplit(moniker, delim)[[1]]
  return(paste0(c(parts[idx]), collapse = ""))
}

partition_fastq_files_table <- function(directory, sample_idx = 1, barcode_idx = 5, lane_idx = 6, read_idx = 7) {
  dt <- data.table(path = dir(directory, full.names = TRUE, pattern = "\\.fastq\\.gz$"))  
  dt[, `:=`(fastq_basename = basename(.SD[,path]),
            read_pair      = read_pair_from_fastq(.SD[,path], read_idx),
            lane           = lane_from_fastq(.SD[,path], lane_idx),
            barcode        = barcode_from_fastq(.SD[,path], barcode_idx),
            sample_name    = sample_name_from_fastq(.SD[,path], sample_idx),
            size           = library_size_from_fastq(.SD[,path], idx = c(2,3))
            ),
         .SDcols = c('path'),
         by = 1:nrow(dt)]
  dt <- dt[dt[,.(fastq_basename, lane, barcode, read_pair)],
           on=.(lane, barcode, read_pair < read_pair)][!is.na(fastq_basename)]
  dt[, path := NULL]
  dt[, read_pair := NULL]
  colnames(dt) <- c("read_1", "lane", "barcode", "sample_name", "library_size", "read_2")
  dt[, read_1:=file.path(directory, read_1)]
  dt[, read_2:=file.path(directory, read_2)]
  return(dt[,.(sample_name, read_1, read_2, library_size, lane, barcode)])
}

```

## Get list of raw fastq files

```{r, echo = FALSE}
x1 <- partition_fastq_files_table(raw_fastq_directory)
knitr::kable(x1)
```

## Subset
Only align a small number of these.
```{r}
x2 <- x1[library_size == '3-5kb']
knitr::kable(x2)
```

## Data Volume?

```{r}
filestats <- data.table(filename = c(x1$read_1, x1$read_2))
filestats[, size := gdata::humanReadable(file.size(filename))]
filestats[, is_read1 := grepl(x = filename, pattern = "_R1\\.")]
filestats[, fastq := basename(filename)]
knitr::kable(filestats[, .(fastq, size, is_read1)])
```

## Get the read pairs for `bwa` we want the files in read pairs...
## Alignments ammended 23/11
Run the alignments with bwa [1], assigning read group identifiers from the targetsfile.
```{r}
submit_bwa <- function(rg_id       = c(),
                       rg_sample   = c(),
                       threads     = c(),
                       pair1       = c(),
                       pair2       = c(),
                       output_dir  = c(),
                       job.init    = c("asub"),
                       job.opts    = c(),
                       stem        = c()) {
  # prepare output file names and directory
  stem <- file.path(output_dir, rg_id)
  dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
    stopifnot(dir.exists(output_dir))
  genome     <- genome_fasta()
  # fill in the command, and write to disk

  cmd <- glue::glue("bwa mem -t {threads} -R '@RG\\tID:{rg_id}\\tSM:{rg_sample}' {genome} {pair1} {pair2} | samtools sort -l 9 -O bam -o {stem}.bam - && samtools index {stem}.bam")
  tmpf <- tempfile(pattern = "submit_bwa.", fileext = ".sh")
  writeLines(text = cmd, con = tmpf)

  # submit the jobs
  submit <- "{job.init} {job.opts} {tmpf}"
  system(glue(submit), intern = TRUE)
}

submit_bwa(rg_id      = paste("Russell", seq_along(rg_sample), sep = "_"),
           rg_sample  = paste("Russell", x2$lane, x2$barcode, sep = "_"),
           threads    = 8,
           pair1      = x2$read_1,
           pair2      = x2$read_2,
           output_dir = "/workspace/hrtcdf/FSTs/Russellalt",
           job.init   = "asub ",
           job.opts   = "-q normal -R rusage[tmp=2000]",   
           stem       = file.path(output_dir, rg_id))

```
##https://www.ibm.com/support/knowledgecenter/SSWRJV_10.1.0/lsf_admin/res_reqs_job_lsf.html

## Merge the 3 Russell bamfiles.
```{r}
t <- 8
bamfiles       <- paste('Russell_1.bam', 'Russell_2.bam', 'Russell_3.bam', sep = " ")
output         <- 'Russell_merge.bam'
samtools_merge <- glue("samtools merge {output} {bamfiles} -r")
logDir         <- "/powerplant/workspace/hrtcdf/FSTs/Russellalt/logDir"
dir.create(logDir, recursive = TRUE, showWarnings = FALSE)
    stopifnot(dir.exists(logDir))
out_file       <- file.path(logDir, "job.out")
err_file       <- file.path(logDir, "job.err")
bsub_cmd       <- glue("bsub -n {t} -o {out_file} -e {err_file} {samtools_merge}")
system(bsub_cmd)
```
## index so -r chr25 can be called in mpileup. 
```{r}
t <- 8
bamfiles       <- paste('Russell_merge.bam', sep = " ")
output         <- 'Russell_merge.bam.bai'
samtools_index <- glue("samtools index {bamfiles} {output} -c")
logDir         <- "/powerplant/workspace/hrtcdf/FSTs/Russellalt/logDir"
dir.create(logDir, recursive = TRUE, showWarnings = FALSE)
    stopifnot(dir.exists(logDir))
out_file       <- file.path(logDir, "job.out")
err_file       <- file.path(logDir, "job.err")
bsub_cmd       <- glue("bsub -n {t} -o {out_file} -e {err_file} {samtools_index}")
system(bsub_cmd)
```

```{r run-mileup for chr25}
submit_mpileup <- function(reference = "genome", 
                           output = c(), 
                           bamfiles = c(), 
                           job.init = "asub", 
                           job.opts = "-q normal -n 8")
{
  cmd <- "bcftools mpileup -B -I -Q 0 -a {mpileuptags} -f {reference} {bamfile} | bcftools call -f {bcfcalltags} -p {pvalThres} -m -A -Oz -o {output} && bcftools index {output}"
  tmpf <- tempfile(pattern = "submit_mileup.", fileext = ".sh")
  writeLines(text = c(glue(cmd)), con = tmpf)
  submit <- "{job.init} {job.opts} {tmpf}"
  system(glue(submit), intern = TRUE)
}
genome      <- genome_fasta()
mpileuptags <- "DP,AD,INFO/AD"
bcfcalltags <- "GQ,GP"
  pvalThres <- 0.99

bamfile     <- "Russell_merge.bam"

out         <- "Russellpile.vcf.gz"
reference   <- dir("/workspace/hrards/projects/hrtcdf/genome/fasta", pattern = ".*\\.gz$", full.names = TRUE)
#file.exists(bamfile)
#file.exists(reference)
submit_mpileup(reference = reference, output = out, bamfiles = bamfiles)
```

```{r filtering for chr25}
bamfiles       <- 'Russellpile.vcf.gz'
output         <- 'Russellpile_chr25.vcf.gz'
logDir         <- "/powerplant/workspace/hrtcdf/FSTs/Russellalt/logDir"

bcftools_filter_chr25 <- glue("bcftools filter -r chr25 -e 'INFO/DP<20 || INFO/DP>200 || QUAL<=20' {bamfiles} -Oz -o {output}")

out_file       <- file.path(logDir, "job.out"); err_file  <- file.path(logDir, "job.err")
bsub_cmd       <- glue("bsub -n 8 -o {out_file} -e {err_file} {bcftools_filter_chr25}")
system(bsub_cmd)
```

```{r additional filtering to only retrieve hets}
bamfiles       <- 'Russellpile.vcf.gz'
output         <- 'Russellpile_chr25alts.vcf.gz'
logDir         <- "/powerplant/workspace/hrtcdf/FSTs/Russellalt/logDir"

bcftools_filteralts <- glue("bcftools filter -r chr25 -e 'INFO/AD[0]=INFO/AD[1] || INFO/DP<30 || INFO/DP>120 || QUAL<=20 || INFO/INDEL=1' -g9 -G9 {bamfiles}  -Oz -o {output}")

out_file       <- file.path(logDir, "job.out"); err_file  <- file.path(logDir, "job.err")
bsub_cmd       <- glue("bsub -n 8 -o {out_file} -e {err_file} {bcftools_filteralts}")
system(bsub_cmd)
```

```{r index Russel2a alts}
#bamfiles       <- 'Russellpile_chr25alts.vcf.gz'
bamfiles       <- 'Russellpile_chr25.vcf.gz'
cmd_index      <- glue("bcftools index -f {bamfiles}")
out_file       <- file.path(logDir, "job.out"); err_file <- file.path(logDir, "job.err")
bsub_cmd       <- glue("bsub -n 8 -o {out_file} -e {err_file} {cmd_index}")
system(bsub_cmd)
```

**finished here moving to merge with male pool vcf's**

#test length - message(length(pair2), pair2)

##to remove temp files from server bsub -m aklppb46  'rm -rf /tmp/*'